---
title: "DM_Project4"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load library
```{r}
library(cluster)
```


Set seed so that the result can be reproducible
```{r}
set.seed(0)
```

Load data to dataframe
```{r}
ff = data.frame(faithful)
```
Scale the data so that the mean is at 0 and sd at 1
```{r}
ff = scale(ff)
```

##Modeling
Plot the data in a scatter plot
```{r}
plot(ff)
```
From the plot, 2 groups of data can be seen. So the number of clusters is 2. The number of clusters is determine using the scatter plot above.

###K-Mean cluster
Plot K-mean with 2 clusters. The algorithm is run 20 times with random central inilization, the best centroid will be kept.
```{r}
km <- kmeans(ff, centers=2, nstart=20)
plot(ff, col=km$cluster)
points(km$centers[,1],km$centers[,2],pch=16,col = 5)
km
```


###Hierarchical using Single link 
```{r}
clusDist = dist(ff)
clusSinLink = hclust(d = clusDist,method = 'single')
```

###Hierarchical using Single link 
```{r}

```






